---
title: "Data Structures in R"
output: html_document
date: "2025-09-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(conflicted) 
library(tidyverse)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
covid_df<-read_csv('covid19.csv')
problems(covid_df)
```

# Positive Covid-19 Cases

## Introduction

(What the project is about) (What your goal is in this project) This project will analyze data collected from different provinces of several countries to determine which country reports the highest number of positive cases in reference to the number of tests conducted

## Read Covid table

```{r}
dim(covid_df)
vector_cols<-colnames(covid_df)
print(vector_cols)
head(covid_df)
glimpse(covid_df)
covid_df
```

## Value of Glimpse Function

The glimpse function is particularly useful since it displays all columns with their respective data types and gives some examples of the data in the cells

```{r}
covid_df_all_states<-covid_df%>%filter(Province_State=='All States')%>%select(-Province_State)
print(covid_df_all_states)
```

## Reason to eliminate province state column

'r (Province_State)' column is removed since all values in that column are "All States" after filtration making it redundant

```{r}
covid_df_all_states_daily<-covid_df_all_states%>%select(c(Date,Country_Region,active,hospitalizedCurr,daily_tested,daily_positive))
covid_df_all_states_daily_sum<-covid_df_all_states_daily%>%group_by(Country_Region)%>%summarize(tested=sum(daily_tested),positive=sum(daily_positive),active=sum(active),hospitalized=sum(hospitalizedCurr))%>%arrange(-tested)

covid_top_10<-head(covid_df_all_states_daily_sum,10)

```

## Display top 10 covid

```{r}
print(covid_top_10)
```

Now we have extracted the top 10 countries have had the highest number of positive against the number of tests \## Name Countries

```{r}
countries<-covid_top_10$Country_Region
tested_cases<-covid_top_10$tested
positive_cases<-covid_top_10$positive
active_cases<-covid_top_10$active
hospitalized_cases<-covid_top_10$hospitalized
names(positive_cases)<-countries
names(tested_cases)<-countries
names(active_cases)<-countries
names(hospitalized_cases)<-countries
```

## Top 3

To determine top 3 positive against tested cases

```{r}
positive_tested<-positive_cases/tested_cases
print(positive_tested)
```

From the output above, we can determine the top three positive against test cases

```{r}
positive_tested_top_3<-c('United Kingdom'= 0.113260617,'United States'= 0.108618191,'Turkey'= 0.080711720)
```

To create matrix with united_kingdom, united_states, turkey

```{r}
united_kingdom<-c(0.11,1473672,166909,0,0)
united_states<-c(0.10,17282363,1877179,0,0)
turkey<-c(0.08,2031192,163941,2980960,0)
covid_mat<-rbind(united_kingdom,united_states,turkey)
colnames(covid_mat)<-c('Ratio','tested','positive','active','hospitalized')
print(covid_mat)
```

## Storing all data for future reference

Storing all the data

```{r}
question<-'Which countries have had the highest number of positive cases against the number of tests?'
answer<-c('Positive tested cases'=positive_tested_top_3)
data_frames<-list(covid_df,covid_df_all_states,covid_df_all_states_daily,covid_df_all_states_daily_sum,covid_top_10)
matrices<-list(covid_mat)
vectors<-list(countries,vector_cols)
data_structure_list<-list("dataframes"=data_frames,"matrices"=matrices,"vectors"=vectors)
covid_analysis_list<-list(question,answer,data_structure_list)
covid_analysis_list[[2]]
```
